MSstatsPTM statistical relative quantification of post-translational modifications in bottom-up proteomic experiments

### 2022-03-21 main/MSstatsPTM_Main.tex
Results - beginning

We propose a statistical method for detecting changes in PTMs while removing confounding with changes in global protein abundance. Here the method is applied to a balanced design with multiple conditions and one technical replicate. In practice the experimental design can be unbalanced and variation can come from multiple sources, as shown in the next section.

Detection of differentially modified PTM sites is performed through testing the null hypothesis of ‘no change’ against the alternative. The null hypothesis states that there is no difference in log-abundance of the PTM site between conditions, adjusted with respect to protein abundance. The expected values of log-abundances of the PTM and protein in Condition $i$ are denoted by $\mu_{i}^{\ast}$ and $\mu_{i}$ respectively. The feature and run level summarization for condition $i$ is denoted as $\hat{y}_{i+}^{\ast}$ for the modified peptide and $\hat{y}_{i+}$ for the unmodified peptide. For two conditions $i$ and $i'$ the hypothesis test is formulated as follows:

>>>>

We propose a statistical method for detecting changes in PTMs while removing confounding with changes in global protein abundance. For simplicity, we discuss the method in the special case of a balanced design with multiple conditions and one technical replicate. However, in  practice the experimental design can be unbalanced and variation can come from multiple sources, as shown in the next section.

\todo{Put it in the context of the figure again, and also in the context of the existing methods in Background - how is it similar to the existing methods, and how different} The null hypothesis states that there is no difference in log-abundance of the PTM site between conditions, adjusted with respect to protein abundance. The expected values of log-abundances of the PTM and protein in Condition $i$ are denoted by $\mu_{i}^{\ast}$ and $\mu_{i}$ respectively. The feature and run level summarization \todo{Summarized how?} for condition $i$ is denoted as $\hat{y}_{i+}^{\ast}$ for the modified peptide and $\hat{y}_{i+}$ for the unmodified peptide. For two conditions $i$ and $i'$ the hypothesis test is formulated as follows:
